---
title: "PASSNYC"
author: "Arjun Chattoraj"
date: "July 13, 2018"
output: html_document
---

# Introduction

This is a dataset available on **PASSNYC** available on *Kaggle*.
The data set is titled: "PASSNYC: Data Science for Good Challenge"
The link to the dataset is: https://www.kaggle.com/passnyc/data-science-for-good/kernels

I want to work on this dataset because it seems like a fun challenge and a great learning experience.

Instead of focusing on Machine Learning and model building,
I want to try my hands on data visualization techniques.

As an up and coming statistician and a data scientist, it is essential to improve in visualization.
It can help in making models efficient, by establishing patterns early, 
instead of relying on the computer to make potentially erroneous errors, 
and removing redundant variables from consideration.

It also looks pretty!

# Packages

```{r, message = F, warning = F}
library(tidyverse)
#devtools::install_github('hadley/ggplot2')
library(ggplot2)
#devtools::install_github("rstudio/leaflet")
library(leaflet)
library(leaflet.minicharts)
#library(corrplot)
#devtools::install_github("kassambara/ggcorrplot")
library(ggcorrplot)
library(plotly)
library(data.table)
library(factoextra)
```


# Loading the data

The working directory is set to where the data files are located.

```{r}
school <- read.csv("2016 School Explorer.csv")
shsat <- read.csv("D5 SHSAT Registrations and Testers.csv")
```

# Data management

There's a few things that are set as factors that need to be numeric

```{r}
school = school[,-c(1:3)]

school[, sapply(school, class) == 'factor'] <- 
    as.data.frame(apply(school[, sapply(school, class) == 'factor'],2, 
                        function(x) gsub("[%$,]", "", x)), stringsAsFactors = FALSE)

school[,c(14:24,26,28,30,32,34,37,38)] <- as.numeric(unlist(school[,c(14:24,26,28,30,32,34,37,38)]))
```


```{r,include = FALSE}
glimpse(school)
```

```{r, include = F}
glimpse(shsat)
```

# Data visualization

## Using a map

```{r}
pal1 = colorNumeric(palette =  "Reds", domain = school$Economic.Need.Index)
econ.index = leaflet(data = school) %>% 
    addTiles() %>%
    addProviderTiles(providers$Stamen.TonerLite) %>%
    addCircles(~Longitude, ~Latitude, radius=~Economic.Need.Index*400, 
               color = ~pal1(Economic.Need.Index), stroke = TRUE, fillOpacity = 1.0,
               popup = paste("City:", school$City, "<br>","School:", school$School.Name, "<br>",
                             "Need Index:", school$Economic.Need.Index),
               group = "Economic.Need.Index") %>%
    addLegend("bottomright", pal = pal1, values = ~Economic.Need.Index, 
              title = "Economic.Need.Index", opacity = 1) %>%
    setView(mean(school$Longitude), mean(school$Latitude), zoom = 10)

econ.index
    
```

```{r}
pal2 = colorNumeric(palette =  "PRGn", domain = school$School.Income.Estimate)
income.est = leaflet(data = school) %>% 
    addTiles() %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addCircles(~Longitude, ~Latitude,  
               color = ~pal2(School.Income.Estimate), stroke = TRUE, fillOpacity = 1.0,
               popup = paste("City:", school$City, "<br>","School:", school$School.Name, "<br>"),
               group = "School.Income.Estimate") %>%
    addLegend("bottomright", pal = pal2, values = ~School.Income.Estimate, 
              title = "School.Income.Estimate", opacity = 1) %>%
    setView(mean(school$Longitude), mean(school$Latitude), zoom = 10)

income.est
 
```

```{r}
pal3 = colorNumeric(palette =  "RdYlBu", domain = school$Student.Attendance.Rate)
att.rate = leaflet(data = school) %>% 
    addTiles() %>%
    addProviderTiles(providers$Stamen.Terrain) %>%
    addCircles(~Longitude, ~Latitude,  
               color = ~pal3(Student.Attendance.Rate), stroke = TRUE, fillOpacity = 1.0,
               popup = paste("City:", school$City, "<br>","School:", school$School.Name, "<br>"),
               group = "Student.Attendance.Rate") %>%
    addLegend("bottomright", pal = pal3, values = ~Student.Attendance.Rate, 
              title = "Student.Attendance.Rate", opacity = 1) %>%
    setView(mean(school$Longitude), mean(school$Latitude), zoom = 10)

att.rate
 
```

It makes sense that lower income schools would have a higher need, as is shown in the graph below.
But the maps can help us see where.
They also help us notice how the data matches with location.

Perhaps, we could try using countour maps or heatmaps to see location-based patterns.

# Graphs
How many of these schools are community Schools?

```{r}
isComm = sum(school$Community.School. == "Yes")
notComm = sum(school$Community.School. == "No")

comm.dt = data.frame(Type = c("Community", "Not Community"), 
                     value = c(isComm*100/(isComm+notComm),notComm*100/(isComm+notComm)))

comm.pie = ggplot(comm.dt, aes(x="", y=value, fill=Type))+
    geom_bar(width = 1,stat="identity") +
    coord_polar("y",start = 0) +
    #scale_fill_manual(values = c("#00BFC4","#F8766D"))+
    theme_light() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks = element_blank()) +
    geom_text(aes(y = value/2),label = paste0(round(comm.dt$value, digits = 2),"%")) +
    ggtitle("Proportion of Community Schools")

comm.pie
```

***

The ggplot2 version of the pie chart is honestly kind of "hacky". 
It takes a bar plot and reorganizes it.
This does not translate well when using ggplotly.
However, plotly has its own pie chart method, 
which is much more intuitive than the ggplot2 method.

Just for kicks, I'm going to recreate the same plot using plotly. It's nicer.

```{r}
school %>%
    group_by(Community.School.) %>%
    summarize(count = n()) %>%
    plot_ly(labels = ~c("Not Community","Community"),
            values = ~count,
            type = "pie",
            marker = list(colors = c("#00BFC4","#F8766D"))) %>%
    layout(title = "Ratio of Community vs Non-Community Schools",
           showlegend = T)
```

That's nicer.

***

```{r}
text.IvN = paste("City:", school$City, "<br>","School:", school$School.Name, "<br>")

inc.v.need = ggplot(school, aes(School.Income.Estimate, Economic.Need.Index, color = Community.School.,
                                text = text.IvN)) + 
    geom_point(size = 0.7)+
    theme_light()+
    guides(color = guide_legend(title="Community School?"))+
    xlab("School Income Estimate")+
    ylab("Economic Need Index")+
    ggtitle("Income Estimate vs Need Index")

ggplotly(inc.v.need)
```

***
A quick comment on `ggplotly`, because it takes a ggplot2 graph and makes it a plotly graph,
there are a few things I would like to change but cant because of the differences between the two.
It's handy to be able to do so, but as a note, I'll stick to using only one from the future,
probably plotly.
***

Let's plot a few other things. 

It'd be interesting to see a correlation plot of the numeric data (not the Grade-wise ones though)

```{r}
school.numeric = school %>%
                    select(Latitude, Longitude, Economic.Need.Index, School.Income.Estimate,
                    Percent.ELL, Percent.Asian, Percent.Black, Percent.Hispanic, 
                    Percent.Black...Hispanic, Percent.White, Student.Attendance.Rate, 
                    Percent.of.Students.Chronically.Absent, Rigorous.Instruction.., 
                    Collaborative.Teachers.., Supportive.Environment.., Effective.School.Leadership..,
                    Strong.Family.Community.Ties.., Trust.., Average.ELA.Proficiency, Average.Math.Proficiency)

corr.mat = cor(school.numeric[complete.cases(school.numeric),])

ggplotly(ggcorrplot(corr.mat, title = "Correlation Matrix", tl.cex = 7))
```

***
In reference to my comment above about using ggplotly,
y'know,
it has it's uses.
This correlation plot was cool, but adding text on top was a bit much.
Adding the interactive features of plotly is handy.
***

Okay, more things. How many schools per city?

```{r}
school %>%
    group_by(City) %>%
    summarize(count = n()) %>%
    plot_ly(labels = ~City,
            values = ~count,
            type = "pie") %>%
    layout(title = "Number of schools per city")
```

Okay, addressing ethinicity.

I wanted to make a map object that produces an ethnicity ratio when you click on a school.
That seems to be a little too complicated.

One way of presenting this is doing city-wise ethnicity comparisons.

```{r}
ethnic.city = school %>%
    gather(c(16:19,21), key = "Ethnicity", value = "Total") %>%
    group_by(City, Ethnicity) %>% 
    na.omit() %>%
    summarise(Percent = round(mean(Total),2)) %>%
    
    setDT() %>%
    dcast(City ~ Ethnicity, value.var = "Percent") 


ethnic.city %>%
    plot_ly(x = ~Percent.Asian, y = ~City, type = 'bar', orientation = 'h', name = "%Asian") %>%
    add_trace(x = ~Percent.Black, name = "%Black") %>%
   # add_trace(x = ~Percent.ELL, name = "%ELL") %>%
    add_trace(x = ~Percent.Hispanic, name = "%Hispanic") %>%
    add_trace(x = ~Percent.White, name = "%White") %>%
    
    layout(barmode = 'stack', xaxis = list(title = ""), yaxis = list(title = "", title))

```

Okay, I don't understand what the problem is bu certain cities' data are not present. One notable exclusion is the Bronx.
I'm going to redo this in ggplot2

```{r}
ggplotly(school %>%
    gather(c(17:19,21), key = "Ethnicity", value = "Total") %>%
    group_by(City, Ethnicity) %>% 
    na.omit() %>%
    summarise(Percent = round(mean(Total),2)) %>%
    ggplot(aes(City, Percent, fill = Ethnicity)) + 
    geom_col(position = "fill")+
    coord_flip())
    
```

That seems better,

For the ELL, I want to see if there is a relationship between ELL, School Income and Need Index.

```{r}
plot_ly(
    data = school,
    x = ~Economic.Need.Index, 
    y = ~School.Income.Estimate,
    color = ~Percent.ELL,
    size = ~Percent.Black...Hispanic,
    text = ~paste("City:", City, "<br>", 
                  "School:", School.Name, "<br>",
                  "ENI:", Economic.Need.Index, "<br>",
                  "Income:", School.Income.Estimate, "<br>",
                  "%ELL:", Percent.ELL, "%<br>",
                  "%Black or Hispanic", Percent.Black...Hispanic, "%<br>"),
    type = "scatter"
)
```

I can try clustering

```{r}
set.seed(7)
sample.indices = sample(1:nrow(school), 0.10*nrow(school))
s1 = school[sample.indices,]
s1 = s1[!duplicated(s1["School.Name"]),]

s1 = s1[,-c(2:13,15,25,27,29,31,33,35,36,39:158)] %>% mutate_if(is.numeric, scale)
s1[is.na(s1)] = 0

glimpse(s1)
```

```{r}

res = s1[-1]
row.names(res) = s1$School.Name

ggplotly(fviz_dist(get_dist(res, stand = TRUE, method = "pearson"), show_labels = F))

```

```{r}
fviz_nbclust(res, kmeans, method = "gap_stat")
```

```{r}
km.res = kmeans(res, 2, nstart = 25)

fviz_cluster(km.res, data = res,
             ellipse.type = "convex")
```










